What if you drop that? Modeling children's curiosity about physical interactions

### 2020-01-26 paper/paper.Rmd
paper

Our computational theory is inspired by classic descriptive theories that posited that curiosity-driven behavior was the consequence of attentional capture by novel stimuli [e.g., @Berlyne1954,@Fantz1964], as well as theories providing qualitative explanations for curiosity-driven behavior as the result of learners' preferences for stimuli that are "moderately discrepant" in relation to their current knowledge state, and thereby provide opportunities to learn [@Kinney1976]. Recent studies have provided empirical support for this basic idea: XX-old infants are sensitive to stimulus complexity and prefer to look at moderately complex stimuli rather than overly simple or overly complex stimuli; @Kidd2012; @Kidd2014].
According to these theories, curiosity-driven behavior reflects a gap between the learner's knowledge and the state of the world [@Loewenstein1994], and thus they predict that as a learner gains additional knowledge, their preferences will shift toward more complex stimuli [@Dember1957].
Although these theories have provided important qualitative insights, one of their major limitations is that they do not provide precise ways to characterize a learner's current knowledge state, how physical states of the world are represented in their mind, nor how discrepancies between a learner's knowledge and such states are compared.

@@

On the computational modeling side, reinforcement learning (RL) researchers have long realized the value of mechanisms that encourage actions that explore the state space in the absence of an apparent reward (which are often sparse), as actions with uncertain outcomes can result in the discovery of new policies with high expected values [@Schmidhuber2010].
Others have built robot models that utilize a form of curiosity to help build predictive and control models of the world [@Oudeyer2007; @Oudeyer2013], although this work pre-trained models on a separate prediction task before implementing curosity-driven learning.
These models' behavior has not yet been directly compared to human behavior in the same task to reveal whether the curiosity driving their behavior mimics people's selection of causal interventions.

<!-- High-level statement of our approach -->
In this paper we propose a new paradigm that allows us to directly compare children's curiosity about physical interactions with that of an aritifical agent implementing different functional forms of curiosity in a simulated 3D environment, carrying out the same task.
The proposed task, described in detail below, presents learners (children or the agent) with a set of objects from which they can select one to drop in a bin containing a set of target objects, with the potential to create one or more collisions during the drop.
Our approach is to coordinate detailed measurement of stimulus and action selection with the construction of task-performing computational models evaluated on the same task and using the same metric, with the goal of developing a more robust, unified, and testable theory of curiosity's role in cognitive development.

>>>>

Our computational theory is inspired by classic descriptive theories that posited that curiosity-driven behavior was the consequence of attentional capture by novel stimuli [e.g., @Berlyne1954,@Fantz1964], as well as theories providing qualitative explanations for curiosity-driven behavior as the result of learners' preferences for stimuli that are "moderately discrepant" in relation to their current knowledge state, and thereby provide opportunities to learn [@Kinney1976]. Recent studies have provided empirical support for this basic idea: XX-old infants are sensitive to stimulus complexity and prefer to look at moderately complex stimuli rather than overly simple or overly complex stimuli; @Kidd2012; @Kidd2014].
According to these theories, curiosity-driven behavior reflects a gap between the learner's knowledge and the state of the world [@Loewenstein1994], and thus they predict that as a learner gains additional knowledge, their preferences will shift toward more complex stimuli [@Dember1957].
Although these theories have provided important qualitative insights, one of their major limitations is that they do not provide precise ways to characterize a learner's current knowledge state, how physical states of the world are represented in their mind, nor how discrepancies between a learner's knowledge and such states are compared.  
Our approach leverages theory and insights from reinforcement learning (RL), including the value of mechanisms that equip RL agents with "intrinsic motivation" — a drive to explore the state space even when rewards are sparse or absent, and favor actions with uncertain outcomes, because they may result in the discovery of new policies with high expected values [@Schmidhuber2010].
Our modeling approach is resonant with recent work that has instantiated such intrinsic-motivation mechanisms in robots to help them learn robust ways to predict physical events in the world world [@Oudeyer2007; @Oudeyer2013], although this work typically involves pretraining models on a separate physical prediction task before implementing curosity-driven learning, and has not yet been directly compared to human behavior in the same prediction tasks.

@@

<!-- High-level statement of our approach -->
In sum, this paper presents a unified paradigm for advancing theories of curiosity and its role in guiding the development of knowledge about the physical world. We present: (1) a computational theory of intrinsic motivation instantiated in an artificial reinforcement-learning agent that explores a simulated 3D environment; (2) a novel dataset that measures children's actions in a novel physical exploration task; and (3) quantitative comparisons between the actions taken by the artificial agent and children on the same set of tasks. Overall, our approach of coordinating the development of task-performing computational models and detailed measurement of human action selection on the same tasks has the promise to lead to more robust and precise theories of how curiosity guides cognitive development.

<!-- presents learners (children or the agent) with a set of objects from which they can select one to drop in a bin containing a set of target objects, with the potential to create one or more collisions during the drop. -->
